An examination of the high-order dynamic interactions underlying multi-dimensional timeseries data

### 2019-03-20 paper/main.tex
remove duplicate from paragraph data
added back the PCA analyses now using the non-dupicaled data
WIP paper and rerun of analyses and figs

\documentclass[english]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{apacite}
\usepackage{natbib}
\usepackage{pxfonts}
\usepackage[utf8]{inputenc}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage[left]{lineno}
\usepackage{soul}
\linenumbers

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/methods_fig}
  \caption{\textbf{Examples of time-varying weights.  Each panel
      displays per-timepoint weights at $t = 50$, evaluated for 100
      timepoints ($1, ..., 100$).  a. Uniform weights.} The weights
    are timepoint-invariant; observations at all timepoints are
    weighted equally, and do not change as a function of $t$.  This is
    a special case of weight function that reduces dynamic
    correlations to static correlations. }
  \label{fig:methods_fig}
\end{figure}

@@

\subsubsection*{Timepoint decoding}
Cite HTFA paper, summarize relevant methods.

To explore how higher-order structure varies with stimulus structure
and complexity, we used a previous neuroimaging dataset
~\cite{SimoEtal16}  in which participants listened to an audio recording of a story; 36 particpants listen to an intact version of the story, 18 participants listen to time-scrambled recordings of the same story where paragraphs were scrambled, 36 particpants listen to word-scrambled version and 36 participants lay in rest condition.

Prior work has shown participants share similar neural responses to richly structured stimuli when compared to stimuli with less structure.  To assess whether the moment-by-moment higher order correlations were reliably preserved across participants, we used inter-subject functional connectivity (ISFC) to isolate the time-varying correlational structure (functional connectivity patterns that were specifically driven by the story participants listened to.  Following the analyses conducted by (HTFA)~\cite{MannEtal18}, We first applied HTFA to the fMRI datasets to obtain a time series of 700 node activities for every participant.  We then computed the dynamic weighted ISFC using a gaussian kernel with a width of 5. We then approximated these dynamic correlation using PCA and computed the dynamic weighted ISFC on the approximations.  We repeated this process up to 10th order approximated correlations.

To assess decoding accuracy, we randomly divided participants for each stimulus into two groups. For the zeroth order, we computed the mean factor activity for each group.  For all subsequent orders up to the tenth order, we computed the mean approximated dynamic ISFC of factor activity for each group. We then took weighted combinations of each order activity, for which we optimized by subdividing one the groups. For each group of participants in turn, we compared these activity patterns (using Pearson correlations) to estimate the story times each pattern corresponded to.

@@

\subsubsection*{Neuroimaging dataset~\citep{SimoEtal16}}
Figures: decoding by level, predictions by level

>>>>

\documentclass[english]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{apacite}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{pxfonts}
\usepackage[utf8]{inputenc}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage[left]{lineno}
\usepackage{soul}
\linenumbers

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/methods_fig}
  \caption{\textbf{Computing higher order correlations. } Correlations
  are computed then approximated to the same size as original data.
  This process is repeated to computer }
  \label{fig:methods_fig}
\end{figure}

@@

\subsubsection*{Timepoint decoding}

To explore how higher-order structure varies with stimulus structure
and complexity, we used a previous neuroimaging dataset
~\cite{SimoEtal16}  in which participants listened to an audio recording of a story; 36 particpants listen to an intact version of the story, 17 participants listen to time-scrambled recordings of the same story where paragraphs were scrambled, 36 particpants listen to word-scrambled version and 36 participants lay in rest condition.

Prior work has shown participants share similar neural responses to
richly structured stimuli when compared to stimuli with less
structure.  To assess whether the moment-by-moment higher order
correlations were reliably preserved across participants, we used
inter-subject functional connectivity (ISFC) to isolate the
time-varying correlational structure (functional connectivity patterns
that were specifically driven by the story participants listened to.
Following the analyses conducted by (HTFA)~\cite{MannEtal18}, we first
applied \textit{hierarchical topographic factor analysis} (HTFA) to the fMRI datasets to obtain a time series of 700 node activities for every participant.  We then computed the dynamic weighted ISFC using a gaussian kernel with a width of 5. We then approximated these dynamic correlation using PCA and computed the dynamic weighted ISFC on the approximations.  We repeated this process up to 10th order approximated correlations.

To assess decoding accuracy, we randomly divided participants for each stimulus into training and testing groups. For the zeroth order, we computed the mean factor activity for each group.  For all subsequent orders up to the tenth order, we computed the mean approximated dynamic ISFC of factor activity for each group.  For each group of participants in turn, we compared these activity patterns (using Pearson correlations) to estimate the story times each pattern corresponded to.

Specifically, we asked, for each timepoint: what are the correlations
between the first group's and second group's activity patterns at each
order.  We then took weighted mixture of the correlation matrices at
each order using mixing parameter, $\phi$, where $0 <\phi< 1$ reflects a
weighted mixture of order based decoding. We optimized  $\phi$, by subdividing the training group. We note that the decoding test we used is a conservative in which we count a timepoint label as incorrect if it is not an exact match.

@@

\subsubsection*{Neuroimaging dataset~\citep{SimoEtal16}}
Figures: decoding by level
For our decoding analysis, we used HTFA-derived node activities
~\cite{MannEtal18} from fMRI data collected as participants listened
to an audio recording of a story (intact condition; 36 participants),
listened to time scrambled recordings of the same story (17
participants in the paragraph-scrambled condition listened to the
paragraphs in a randomized order and 36 in the word-scrambled
condition listened to the words in a randomized order), or lay resting
with their eyes open in the scanner (rest condition; 36
participants). We sought to demonstrate how higher-order correlations
may be used to examine dynamic interactions of brain patterns in
(real) multi-subject fMRI datasets. This story listening dataset was
collected as part of a separate study, where the full imaging
parameters, image preprocessing methods, and experimental details may
be found~\citep{SimoEtal16}. The dataset is available at
\href{url}{http://arks. princeton.edu/ark:/88435/dsp015d86p269k}.

We next evaluated if our model of high-order correlations in brain activity can capture cognitively relevant brain patterns. We performed a decoding analysis, using cross validation to estimate (using other participants’ data) which parts of the story each weighted-mixture of higher-order brain activity pattern corresponded to (see \textit{Materials and methods}). We note that our primary goal was not to achieve perfect decoding accuracy, but rather to use decoding accuracy as a benchmark for assessing whether different neural features specifically capture cognitively relevant brain patterns.

Separately for each experimental condition, we divided participants
into two groups. For the zeroth order, we computed the mean factor
activity for each group.  For all subsequent orders up to the tenth
order, we computed the mean approximated dynamic ISFC of factor
activity for each group (see \textit{Materials and methods}).  For
each order, we correlated the group 1 activity patterns with group 2
activity patterns.  We then subdivided the group 1 to obtain an
optimal weighting parameter for each order’s correlation matrix using
the same cross validation method. We used the optimal weighting
parameters to obtain a weighted-mixture of each order’s correlation
matrix. Using these correlations, we labeled the group 1 timepoints
using the group 2 timepoints with which they were most highly
correlated; we then computed the proportion of correctly labeled group
1 timepoints. (We also performed the symmetric analysis whereby we
labeled the group 2 timepoints using the group 1 timepoints as a
template.) We repeated this procedure 10 times (randomly re-assigning
participants to the two groups each time) to obtain a distribution of
decoding accuracies for each experimental condition. (There were 272
timepoints for paragraph condition, 300 timepoints for intact and word
conditions, and 400 timepoints for rest condition,  so chance
performance on this decoding test is was $\frac{1}{272}$,
$\frac{1}{300}$, and $\frac{1}{400}$ respectively.
